Casting doubt on the causal link between intelligence and age at first intercourse:\\ A cross-generational sibling comparison design using the NLSY

### 2016-06-17 Manuscript/Manuscript.Rnw
resubmitted

\documentclass[a4paper,man,apacite,natbib,12pt,longtable,mask]{apa6}
\usepackage{../Common/style/Mason}
\usepackage{listings}
\usepackage{inconsolata}

%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Is the Link Between Intelligence and Age at First Intercourse Causal or Correlational? \\ A Cross-Generational Sibling Comparison Design Using the NLSY}
\shorttitle{Intelligence and AFI}
%
% Authors and Affiliations
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
%
% Author Note
\authornote{\input{../Common/content/title/authornote.tex}}
%
% Abstract
\abstract{%Last compiled \today\ at \currenttime\\
\input{../Common/content/title/abstract_empreport.txt}}
%
% Key Words
\keywords{Age at first intercourse; Cross-sectional Data; Discordant Sibling Design; Intelligence; Quasi-Experimental; Siblings}
%
\begin{document}
\maketitle

@@

  listing <- function(x, options){
    paste("\\begin{lstlisting}[style=Rsettings]\n",
      x, "\\end{lstlisting}\n", sep = "")}
  knit_hooks$set(source=listing, output=listing)
@
<<setup, include=FALSE>>=
# Libraries
## Install Most Current Version
library(devtools)#;dev_mode(on=T)
#install_github("LiveOak/NlsyLinks")
#install_github("harrelfe/Hmisc");dev_mode(on=F)
## Attach Remaining
library(corrplot)
library(ggplot2);library(ggExtra)
library(Hmisc)
library(MplusAutomation)
library(NlsyLinks)
library(plyr);library(psych)
#library(stargazer)
library(tables)
library(xtable)
#
# File Structure
file_drive= "C:"
##file_drive="E:"
directory = paste0(file_drive,"/Dropbox/Lab/Research/Projects/Current/NLSY AFI")
#
# Functions
## source("https://raw.githubusercontent.com/smasongarrison/Functions/master/Mason_Functions.R")
#Modded Stargazer
source(paste0(directory,"/Common/content/functions/stargazer.R"))
source(paste0(directory,"/Common/content/functions/stargazer-internal.R"))
## cor_1star
### Prints out clean correlation matrix. If you want to replace the upper triangle with the sample size, toggle include.n, whose default is FALSE.
cor_1star <- function(x,digit=3,sig=.05,include.n=FALSE){ 
  require(Hmisc) 
  x <- as.matrix(x) 
  R <- rcorr(x)$r 
  p <- rcorr(x)$P 
 ## truncate the matrix that holds the correlations to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), digit))[,-1] 
## build a new matrix that includes the correlations with their appropriate stars 
  Rnew <- R
  if(!is.na(sig)){
## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < sig, "* ", " ")

@@

# Summary Stats by Race and Gender
capture.output(latex(
  tabular((Heading()*( 1 +RowFactor(CSEX)*RowFactor(Race)+ RowFactor(CSEX)+ RowFactor(Race) )~(Heading("AFI")*AFI_G2*(Mean + Sd))), 
          data = Gen2AFIConcor_2),
  numeric.dollar=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_AFIRACEGENDER.tex"))
@
<<setup_tables,include=FALSE,cache=TRUE>>=
# Summary Stats G1
## Manage data
### Import
df <- merge(Gen1AFIConcor,birthweight_fb,by="MCASEID",all=T)
nlsyextract <- read.csv(paste0(directory,"/Common/content/data/nlsyextract.csv"))
###
### Clean
nlsyextract$IDC1[nlsyextract$IDC1 <= 0  ] <- 0
nlsyextract$IDC1[nlsyextract$IDC1 > 0  ] <- 1
nlsyextract$SiblinginSample <-nlsyextract$IDC1
nlsyextract$IDC1 <- NULL
nlsyextract[nlsyextract< 0  ] <- NA
nlsyextract$month<-nlsyextract$Q1.3_A.M 
nlsyextract$month<- paste("0",nlsyextract$Q1.3_A.M,sep="")
nlsyextract$month[nlsyextract$month == 010 ] <- 10
nlsyextract$month[nlsyextract$month == 011 ] <- 11
nlsyextract$month[nlsyextract$month == 012 ] <- 12
nlsyextract$Age<-as.numeric(as.Date(paste(df$YRB,"-06-15",sep=""), format='%Y-%m-%d')-as.Date(paste("19",nlsyextract$Q1.3_A.Y,"-",nlsyextract$month,"-15",sep=""), format='%Y-%m-%d'))/365.25
summarystats_g1<-data.frame(nlsyextract$SiblinginSample,nlsyextract$Age,df$Intel_G1,df$AFI_G1,nlsyextract$SAMPLE_SEX)
###
### Sibling in sample?
#### Gen2AFIConcor$MCASEID<-birthweight$MCASEID
#### data=Gen2AFIConcor
#### data$dup<-0
#### for (x in 1:length(data[,1])) {
#### for (y in 1:length(data[,1])) {
#### if (data$MCASEID[x]==data$MCASEID[y]) data$dup[x]= data$dup[x]+1 }}
#### write.table(data, file = paste0(directory,"/Common/content/data/g2sibinsample.csv"), append = FALSE)
Gen2AFIConcor<- read.table(paste0(directory,"/Common/content/data/g2sibinsample.csv"), header=TRUE, quote="\"")
Gen2AFIConcor$SiblinginSample<-1
Gen2AFIConcor$SiblinginSample[Gen2AFIConcor$dup == 1 ]<- 0
summarystats_g2 <- merge(categories_g2,Gen2AFIConcor,by="CASEID",all=T)
summarystats_g2$zIntel_G2<-scale(summarystats_g2$Intel_G2)
summarystats_g2 <- merge(summarystats_g2,birthweight_2,by="CASEID",all=T)
##
## Capture output
capture.output(latex(
  tabular(((Heading("Mother Age at Birth")*nlsyextract.Age + Heading("Standardized AFQT")*df.Intel_G1 + Heading("AFI")*df.AFI_G1 ) ~ Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*( 1 + RowFactor(nlsyextract.SiblinginSample) )), data = summarystats_g1),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg1.tex"))

@@

%%%%%%%%%%%% Method %%%%%%%%%%%%
\section{Method}
\subsection{Research Design}
\input{../Common/content/writing/design.tex}
%
% Sample
\subsection{Sample}
\input{../Common/content/writing/subjects.tex}
%
% Tetrads
\subsection{Tetrads}
\input{../Common/content/writing/tetrad.tex}
%
%%%%%%%%%%%% Measurement %%%%%%%%%%%%
% Age at First Intercourse
\subsection{Measures}
\input{../Common/content/writing/AFI.tex}
\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Smoothed Density Plots of AFI by Race and Sex}
\label{plot_afi_by_race_sex}
\begin{center}
<<plot_afi_by_race_sex,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.8, fig.width=5,out.width=".65\\paperwidth",warning=FALSE>>=
# Data Prep
dataframe<-data.frame(Gen2AFIConcor_2,Gen2AFIConcor$Intel_G2)
names(dataframe)<- c("CASEID","MCASEID","Race","SEX","AFI","Intel")

@@

ggplot(dataframe, aes(x=AFI, colour=RacebySEX, group=RacebySEX, linetype = RacebySEX))+ geom_density(fill=NA,adjust=2)+ theme_min(size=10)+  
    theme(legend.position="bottom",
          legend.justification=c(1, 0.5),
          legend.text=element_text(size=6),
          legend.key = element_rect(fill = 'white', size = 0.5),
          legend.title = element_blank()) +   
    xlab("AFI") + 
    ylab("Density") +
    scale_colour_manual(values= c("#56B4E9","#56B4E9", "#009E73", "#009E73", "#CC79A7", "#CC79A7")) +
    coord_cartesian(ylim=c(0,0.25), xlim=c(12,25)) +
    scale_linetype_manual(values=c(1,2,1,2,1,2))
  @
\end{center}
\end{minipage}
%
% Intelligence
\input{../Common/content/writing/intelligence.tex}
%
% Difference Score Reliability
\subsection{Reliability of Difference Scores}
<<calculationalphadif,cache=TRUE,include=FALSE>>=
# Intelligence
sigmax = sqrt(1.1881)
sigmay = sigmax
rxx = .87
ryy = rxx
rxy=.67
r.diff.intelligence <- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) /(sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
#AFI
sigmax = sqrt(4.41)
sigmay = sigmax
rxx = .76
ryy = rxx
rxy=.099
r.diff.afi <- (sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) / (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay)
@
Our design assumes that the difference scores of our measures are reliable. We've reported the test-retest reliability of Gen2 intelligence and Gen1 AFI in earlier sections. Here, we report the test-retest reliability of the pairwise differences of those measures. 
\subsubsection{AFI} Comparing sibling differences in AFI as reported in 1983 and 1984 (n = 783 pairs) we found a moderate correlation (r = .76). The sample of sibling pairs with complete information in 1985 was too small (n = 12 pairs) to compare to the other two years. Regardless, sibling differences in self-reported AFI appear satisfactory for research purposes. We note that the common concern with unreliability of difference scores is at least partially mitigated when the separate scores defining the differences are reliable themselves. Because we could not estimate test-retest reliabilities for Generation 2, we calculated the reliability using Lord's equation (\citeyear{Lord1963}). Generation 2 Mean AFI difference scores were also moderately reliable (r = \Sexpr{r.diff.afi}) and comparable to Generation 1 sibling differences.

\subsubsection{Intelligence} Cousin differences in intelligence as assessed at ages 9.5, 10.5, and 11.5 were correlated using three different linking methods (Mixed, Daughters, Sons). Reliabilities across linking methods was consistent and high (min r = .86; max r = .95). However, we could not estimate test-retest reliabilities for Generation 1; we calculated the reliability using Lord's equation (\citeyear{Lord1963}). The calculated reliability of Generation 1's differences in AFQT was .61, which was only minimally  acceptable, and obviously lower than the empirical correlation we derived for cousin differences.%\Sexpr{r.diff.intelligence}

%%%%%%%%%%%% Results %%%%%%%%%%%%
\section{Results}
\input{../Common/content/writing/results.tex}
%
% Between-Family
\subsection{Between-Family Analyses}
\input{../Common/content/writing/results_between.tex}
%
\begin{landscape}\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Generation 2: Age at First Intercourse vs. Intelligence}
\label{plot_gen2_afi}
\begin{center}
<<plot_gen2_afi,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.8, fig.width=5,out.width=".8\\paperwidth",warning=FALSE>>=
library(ggplot2)
library(gtable)
library(grid)
library(gridExtra)
library(NlsyLinks)
# Gen 2 Plot
shading<-c("firebrick4",#"firebrick3","firebrick2",
           "firebrick1",#"mediumvioletred",
             "dodgerblue1",#"dodgerblue2","dodgerblue3",
           "dodgerblue4")
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsOutcomes <- Gen2AFIConcor_3

dsOutcomes$SubjectTag <-dsOutcomes$MCASEID*100
dsOutcomes<-dsOutcomes[!is.na(dsOutcomes$SubjectTag),]
ds2D <-dsOutcomes#
max_val <- max(abs(ds2D$Intel_G1),na.rm=TRUE)
values <- seq(-max_val, max_val, length = length(shading))
##############
ggplot(ds2D, aes(y=AFI_G2, x=Intel_G2,colour=(Intel_G1))) + 
  geom_point(size =1,alpha = 1,na.rm=TRUE,position = position_jitter(w = 0.01, h = 0.5)) +
  scale_color_gradientn(values=values,rescaler = function(x, ...) x, oob = identity,name="Gen1 \nAFQT",colours=shading)+
  theme_min(size=10) +
     theme(plot.title=element_text(size=16,face="bold"))+ 
  ylab("Gen2 Age at First Intercourse") +
  xlab("Gen2 General Cognitive Ability") #+
#ggtitle("Generation 1:\n Age at First Intercourse vs. Intelligence")
@
\end{center}
\end{minipage}\end{landscape}
%
% Within-Family
\subsection{Within-Family Analyses}
\input{../Common/content/writing/results_within.tex}
%
\begin{landscape}
\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Generation 1: Age at First Intercourse vs. Intelligence}
\label{plot_gen1_afi}
\begin{center}
<<plot_gen1_afi,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.8, fig.width=5,out.width=".8\\paperwidth",warning=FALSE>>=
library(ggplot2)
library(gtable)
library(grid)
library(gridExtra)
library(NlsyLinks)
# Gen 1 Plot
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsOutcomes <- Gen1AFIConcor
dsOutcomes$Sex<-testretest$SEX
dsOutcomes$SubjectTag <-dsOutcomes$MCASEID*100
dsD <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c("Intel_G1","AFI_G1","Sex"))
dsD$DIF_G<-#abs
  (dsD$Intel_G1_S1-dsD$Intel_G1_S2)
dsD$DIF_A<-#abs
  (dsD$AFI_G1_S1-dsD$AFI_G1_S2)
#m<-data.frame(Gen2AFIConcor$Intel_G2,Gen2AFIConcor$AFI_G2)

#shading<-c("navy","dodgerblue4","dodgerblue3","dodgerblue2","lightsteelblue1","dodgerblue2","dodgerblue3","dodgerblue4","navy")
shading<-c("firebrick4",#"firebrick3","firebrick2",
           "firebrick1",#"mediumvioletred",
             "dodgerblue1",#"dodgerblue2","dodgerblue3",
           "dodgerblue4")
max_val <- max(abs(dsD$DIF_G),na.rm=TRUE)
values <- seq(-max_val, max_val, length = length(shading))

dsD$density<-NA
dsD$density[dsD$DIF_G < .33]<- 0
dsD$density[dsD$DIF_G < -.33]<- -1
dsD$density[dsD$DIF_G >= .33]<- 1
dsD$density<-as.factor(dsD$density)

### Marginal Density Plot y
ydensity <- ggplot(dsD, aes(x=AFI_G1_S1, group=density, color=density)) + 
  geom_density(fill=NA,adjust=2)+
  scale_colour_manual(values= c("firebrick1", "#AD78B6", "dodgerblue1"))+
  coord_cartesian(ylim=c(0,0.25), xlim=c(12,25))+
  theme_min(size=8)+ 
  theme(legend.position = "none",
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
   plot.margin=unit(c(0,0,.35,-.5), "cm"))+
  coord_flip()
### Marginal Density Plot x

xdensity <- ggplot(dsD, aes(x=Intel_G1_S1, group=density, color=density)) +
  geom_density(fill=NA,adjust=2) +
  scale_colour_manual(values= c("firebrick1", "#AD78B6", "dodgerblue1"))+ 
  theme_min(size=8) +
  theme(legend.position = "none",axis.title.y = element_blank(),
                            axis.title.x = element_blank(),
                            axis.text.x = element_blank(),
   plot.margin=unit(c(0,0,-.5,2.8), "cm"))

### Scatter plot
main<- ggplot(dsD, aes(y=AFI_G1_S1, x=Intel_G1_S1,colour=(DIF_G))) + 
  geom_point(size =1,alpha = 1,na.rm=TRUE,position = position_jitter(w = 0.1, h = 0.5)) +
  scale_colour_gradientn(name="Gen1 \nSibling \nDifferences \nin AFQT",colours=shading,values=values,rescaler = function(x, ...) x, oob = identity)+
  theme_min(size=10) +
     theme(plot.title=element_text(size=12),
   plot.margin=unit(c(0,0.01,0,-.5), "cm"),
   legend.position="left",
   legend.key.width=unit(.5,"cm"))+ 
  ylab("Gen1 Age at First Intercourse") +
  xlab("Gen1 AFQT (Z Score)") #+
#ggtitle("Generation 1:\n Age at First Intercourse vs. Intelligence")

>>>>

\documentclass[a4paper,man,apacite,natbib,12pt,longtable,mask]{apa6}
\usepackage{../Common/style/Mason}
\usepackage{listings}
\usepackage{inconsolata}
\usepackage[bottom]{footmisc} %fix footnotes
\usepackage{lineno}
\usepackage{adjustbox}
%\linenumbers
\interfootnotelinepenalty=10000
%%%%%%%%%%%% Title %%%%%%%%%%%%
\title{Casting doubt on the causal link between intelligence and age at first intercourse:\\ A cross-generational sibling comparison design using the NLSY}
\shorttitle{Intelligence and AFI}
%
% Authors and Affiliations
\author{S. Mason Garrison and Joseph Lee Rodgers}
\affiliation{Vanderbilt University}
%
% Author Note
\authornote{\input{../Common/content/title/authornote.tex}}
%
% Abstract
\abstract{%Last compiled \today\ at \currenttime\\
\input{../Common/content/title/abstract_empreport.txt}}
%
% Key Words
\keywords{Age at first intercourse; Cross-sectional Data; Discordant Sibling Design; Intelligence; Quasi-Experimental; Siblings}
%
\begin{document}
\maketitle

@@

  listing <- function(x, options){
    paste("\\begin{lstlisting}[style=Rsettings]\n",
      x, "\\end{lstlisting}\n", sep = "")}
  knit_hooks$set(source=listing, output=listing)
@
<<setup, include=FALSE>>=
# Libraries
## Install Most Current Version
library(devtools)#;dev_mode(on=T)
#install_github("LiveOak/NlsyLinks")
#install_github("harrelfe/Hmisc");dev_mode(on=F)
## Attach Remaining
library(corrplot)
library(ggplot2);library(ggExtra);library(gtable)
library(grid);library(gridExtra)
library(Hmisc)
library(MplusAutomation)
library(NlsyLinks)
library(plyr);library(psych)
library(QuantPsyc)
#library(stargazer)
library(tables)
library(xtable)

#
# File Structure
file_drive= "C:"
file_drive="E:"
directory = paste0(file_drive,"/Dropbox/Lab/Research/Projects/Completed/NLSY AFI")
#
# Functions
## source("https://raw.githubusercontent.com/smasongarrison/Functions/master/Mason_Functions.R")
#Modded Stargazer
source(paste0(directory,"/Common/content/functions/stargazer.R"))
source(paste0(directory,"/Common/content/functions/stargazer-internal.R"))
## cor_1star
### Prints out clean correlation matrix. If you want to replace the upper triangle with the sample size, toggle include.n, whose default is FALSE.
cor_1star <- function(x,digit=3,sig=.05,include.n=FALSE){ 
  require(Hmisc) 
  x <- as.matrix(x) 
  R <- rcorr(x)$r 
  p <- rcorr(x)$P 
 ## truncate the matrix that holds the correlations to two decimal
  R <- format(round(cbind(rep(-1.11, ncol(x)), R), digit))[,-1] 
## build a new matrix that includes the correlations with their appropriate stars 
  Rnew <- R
  if(!is.na(sig)){
## define notions for significance levels; spacing is important.
  mystars <- ifelse(p < sig, "* ", " ")

@@

# Summary Stats by Race and Gender
capture.output(latex(
  tabular((Heading()*( 1 +RowFactor(CSEX)*RowFactor(Race)+ RowFactor(CSEX)+ RowFactor(Race) )~(Heading("AFI")*AFI_G2*(Mean + Sd))), 
          data = Gen2AFIConcor_2),
  numeric.dollar=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_AFIRACEGENDER.tex"))
@
<<setup_tables,include=FALSE,cache=TRUE>>=
# Summary Stats G1
## Manage data
### Import
df <- merge(Gen1AFIConcor,birthweight_fb,by="MCASEID",all=T)
nlsyextract <- read.csv(paste0(directory,"/Common/content/data/nlsyextract.csv"))
###
### Clean
nlsyextract$IDC1[nlsyextract$IDC1 <= 0  ] <- 0
nlsyextract$IDC1[nlsyextract$IDC1 > 0  ] <- 1
nlsyextract$SiblinginSample <-nlsyextract$IDC1
nlsyextract$IDC1 <- NULL
nlsyextract[nlsyextract< 0  ] <- NA
nlsyextract$month<-nlsyextract$Q1.3_A.M 
nlsyextract$month<- paste("0",nlsyextract$Q1.3_A.M,sep="")
nlsyextract$month[nlsyextract$month == 010 ] <- 10
nlsyextract$month[nlsyextract$month == 011 ] <- 11
nlsyextract$month[nlsyextract$month == 012 ] <- 12
nlsyextract$Age<-as.numeric(as.Date(paste(df$YRB,"-06-15",sep=""), format='%Y-%m-%d')-as.Date(paste("19",nlsyextract$Q1.3_A.Y,"-",nlsyextract$month,"-15",sep=""), format='%Y-%m-%d'))/365.25
summarystats_g1<-data.frame(nlsyextract$SiblinginSample,nlsyextract$Age,df$Intel_G1,df$AFI_G1,nlsyextract$SAMPLE_SEX)
###
### Sibling in sample?
#### Gen2AFIConcor$MCASEID<-birthweight$MCASEID
#### data=Gen2AFIConcor
#### data$dup<-0
#### for (x in 1:length(data[,1])) {
#### for (y in 1:length(data[,1])) {
#### if (data$MCASEID[x]==data$MCASEID[y]) data$dup[x]= data$dup[x]+1 }}
#### write.table(data, file = paste0(directory,"/Common/content/data/g2sibinsample.csv"), append = FALSE)
Gen2AFIConcor<- read.table(paste0(directory,"/Common/content/data/g2sibinsample.csv"), header=TRUE, quote="\"")
Gen2AFIConcor$SiblinginSample<-1
Gen2AFIConcor$SiblinginSample[Gen2AFIConcor$dup == 1 ]<- 0
summarystats_g2 <- merge(categories_g2,Gen2AFIConcor,by="CASEID",all=T)
summarystats_g2$zIntel_G2<-scale(summarystats_g2$Intel_G2)
summarystats_g2 <- merge(summarystats_g2,birthweight_2,by="CASEID",all=T)
summarystats_g2$YRB[summarystats_g2$YRB<0]<-NA
library(plyr)
x<-count(summarystats_g2$YRB)
z<-sum(x$freq[x$x>1996],na.rm=TRUE)
x<-NULL
z<-NULL
##
## Capture output
capture.output(latex(
  tabular(((Heading("Mother Age at Birth")*nlsyextract.Age + Heading("Standardized AFQT")*df.Intel_G1 + Heading("AFI")*df.AFI_G1 ) ~ Heading("Has Sibling in NLSY?")*(Mean + Sd)*Heading()*( 1 + RowFactor(nlsyextract.SiblinginSample) )), data = summarystats_g1),
  numeric.dollar=TRUE,na.blank=TRUE),
  file=paste0(directory,"/Common/content/tables/table_summary_stats_sibinsampleg1.tex"))

@@

%%%%%%%%%%%% Method %%%%%%%%%%%%
\section{Method}
\subsection{Research Design and Analytic Approach}
\input{../Common/content/writing/design.tex}
%
% Sample
\subsection{Sample}
\input{../Common/content/writing/subjects.tex}
%
% Tetrads
\subsection{Tetrads}
\input{../Common/content/writing/tetrad.tex}
%
%%%%%%%%%%%% Measurement %%%%%%%%%%%%
% Age at First Intercourse
\subsection{Measures}
\input{../Common/content/writing/AFI.tex}
\noindent\begin{minipage}{\linewidth}
\captionof{figure}{Smoothed Density Plots of Gen1 AFI by Race and Sex}
\label{plot_afi_by_race_sex}
\begin{center}
<<plot_afi_by_race_sex,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.8, fig.width=5,out.width=".65\\paperwidth",warning=FALSE>>=
# Data Prep
dataframe<-data.frame(Gen2AFIConcor_2,Gen2AFIConcor$Intel_G2)
names(dataframe)<- c("CASEID","MCASEID","Race","SEX","AFI","Intel")

@@

ggplot(dataframe, aes(x=AFI, colour=RacebySEX, group=RacebySEX, linetype = RacebySEX))+ geom_density(fill=NA,adjust=2)+ theme_min(size=10)+  
    theme(legend.position="bottom",
          legend.justification=c(1, 0.5),
          legend.text=element_text(size=6),
          legend.key = element_rect(fill = 'white', size = 0.5),
          legend.title = element_blank()) +   
    xlab("AFI") + 
    ylab("Density") +
    scale_colour_manual(values= c("#56B4E9","#56B4E9", "#009E73", "#009E73", "#CC79A7", "#CC79A7")) +
    coord_cartesian(ylim=c(0,0.25), xlim=c(12,25)) +
    scale_linetype_manual(values=c(1,2,1,2,1,2))
  @
\end{center}
\end{minipage}
%
% Intelligence
\input{../Common/content/writing/intelligence.tex}
%
% Difference Score Reliability
\subsection{Reliability of Difference Scores}
<<calculationalphadif,cache=TRUE,include=FALSE>>=
# Intelligence
sigmax = sqrt(1.1881)
sigmay = sigmax
rxx = .9#.87
ryy = rxx
rxy=.67
(r.diff.intelligence <- round((sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) /(sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay),digits = 2))
#AFI
sigmax = sqrt(4.41)
sigmay = sigmax
rxx = .76
ryy = rxx
rxx.afi=rxx
rxy=.099
(r.diff.afi <- round((sigmax*sigmax*rxx +sigmay*sigmay*ryy -2*rxy*sigmax*sigmay) / (sigmax*sigmax + sigmay*sigmay -2*rxy*sigmax*sigmay),digits = 2))

@
Our design assumes that the difference scores of our measures are reliable. We have reported the test-retest reliability of Gen2 intelligence and Gen1 AFI in earlier sections. Here, we report the test-retest reliability of the pairwise differences of those measures. 
\subsubsection{AFI} Comparing sibling differences in AFI as reported in 1983 and 1984 (n = 783 pairs) we found a moderate correlation (r = \Sexpr{rxx.afi}). The sample of sibling pairs with complete information in 1985 was too small (n = 12 pairs) to compare to the other two years. Regardless, sibling differences in self-reported AFI appear satisfactory for research purposes. We note that the common concern with unreliability of difference scores is at least partially mitigated when the separate scores defining the differences are reliable themselves. Because we could not estimate test-retest reliabilities for Generation 2, we calculated the reliability using Lord's (\citeyear{Lord1963}) equation. Generation 2 Mean AFI difference scores were also moderately reliable (r = \Sexpr{r.diff.afi}) and comparable to Generation 1 sibling differences.

\subsubsection{Intelligence} Cousin differences in intelligence as assessed at ages 9.5, 10.5, and 11.5 were correlated using three different linking methods (Mixed, Daughters, Sons). Reliabilities across linking methods was consistent and high (min r = .86; max r = .95). However, we could not estimate test-retest reliabilities for Generation 1; we calculated the reliability using Lord's (\citeyear{Lord1963}) equation. The calculated reliability of Generation 1's differences in AFQT was \Sexpr{r.diff.intelligence}0, which was marginally acceptable, and obviously lower than the empirical correlation we derived for cousin differences.%\Sexpr{r.diff.intelligence}

%%%%%%%%%%%% Results %%%%%%%%%%%%
\section{Results}
\input{../Common/content/writing/results.tex}
%
% Between-Family
\subsection{Between-Family Analyses}
\input{../Common/content/writing/results_between.tex}

\noindent\begin{minipage}{\paperwidth}
\captionof{figure}{Gen2 AFI vs. Gen1 \& Gen2 Intelligence}
\label{plot_gen2_afi}
\begin{center}
<<plot_gen2_afi,fig=TRUE,echo=FALSE,cache=TRUE,echo=FALSE, fig.height=5*.6, fig.width=5,out.width="1\\paperwidth",warning=FALSE>>=

substrRight <- function(x, n){
  substr(x, nchar(x)-n+1, nchar(x))
}
# Gen 2 Plot
shading<-c("firebrick4",#"firebrick3","firebrick2",
           "firebrick1",#"mediumvioletred",
             "dodgerblue1",#"dodgerblue2","dodgerblue3",
           "dodgerblue4")
dsOutcomes <- Gen2AFIConcor_3

## Sibling Difference Score
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen2Siblings")

# Source https://amywhiteheadresearch.wordpress.com/2013/01/22/randomly-deleting-duplicate-rows-from-a-dataframe-2/
duplicated.random = function(x, incomparables = FALSE, ...) 
{ 
     if ( is.vector(x) ) 
     { 
         permutation = sample(length(x)) 
         x.perm      = x[permutation] 
         result.perm = duplicated(x.perm, incomparables, ...) 
         result      = result.perm[order(permutation)] 
         return(result) 
     } 
     else if ( is.matrix(x) ) 
     { 
         permutation = sample(nrow(x)) 
         x.perm      = x[permutation,] 
         result.perm = duplicated(x.perm, incomparables, ...) 
         result      = result.perm[order(permutation)] 
         return(result) 
     } 
     else 
     { 
         stop(paste("duplicated.random() only supports vectors", 
                "matrices for now.")) 
     } 
}

dsOutcomes$ASEID<-as.numeric(substrRight(as.character(dsOutcomes$CASEID),1))
dsOutcomes$SubjectTag <-dsOutcomes$CASEID
dsOutcomes<-dsOutcomes[!is.na(dsOutcomes$SubjectTag),]
ds2D <-dsOutcomes#
v<-dsOutcomes[dsOutcomes$ASEID<3,]

ds22D <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c("Intel_G2","AFI_G2"))
ds22D$DIF_G<-#abs
  (ds22D$Intel_G2_S1-ds22D$Intel_G2_S2)
ds22D$DIF_A<-#abs
  (ds22D$AFI_G2_S1-ds22D$AFI_G2_S2)
x<-data.frame(ds22D[,c(1,10,11)])
x<-x[!is.na(x$DIF_G | x$DIF_A),]
x$duplicated <- duplicated.random(x$SubjectTag_S1)
x<- x[x$duplicated==FALSE,]
x$duplicated<-NULL
ds2D <- merge(ds2D,x,by.x="SubjectTag",by.y="SubjectTag_S1",all = TRUE,sort=FALSE)
x<-NULL
v<-NULL
ds22D<-NULL

####
coef<-coef(lm(AFI_G2 ~ Intel_G2, data = dsOutcomes))
max_val <- max(abs(ds2D$DIF_G),na.rm=TRUE)
values <- seq(-max_val, max_val, length = length(shading))
ds2D<-ds2D[!is.na(ds2D$DIF_G),]
##############
g2<-ggplot(ds2D, aes(y=AFI_G2, x=Intel_G2,colour=(DIF_G))) + 
  geom_point(size =.5,alpha = 1,na.rm=TRUE,position = position_jitter(w = 0.015, h = 1)) + geom_abline(intercept = coef[1], slope = coef[2])+
  scale_color_gradientn(values=values,rescaler = function(x, ...) x, oob = identity,name="Sibling \nDifferences \nin Ability",colours=shading,na.value = "#AD78B6")+
  theme_min(size=8) +
     theme(plot.title=element_text(size=8),
   plot.margin=unit(c(0,0,0,0.5), "cm"),
   legend.position= "none"
   #,legend.key.width=unit(.5,"cm")
   )+ 
  ylab("Gen2 Age at First Intercourse") +
  xlab("Gen2 General Ability (Age 9.5)")+coord_cartesian(ylim=c(12,30))+
theme(text = element_text(size=8)) #+
#ggtitle("Generation 1:\n Age at First Intercourse vs. Intelligence")

##### g1
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen1Housemates")
dsOutcomes <- Gen1AFIConcor
dsOutcomes$Sex<-testretest$SEX
dsOutcomes$SubjectTag <-dsOutcomes$MCASEID*100

dsD <- CreatePairLinksDoubleEntered(
outcomeDataset = dsOutcomes,
linksPairDataset = dsLinking,
outcomeNames = c("Intel_G1","AFI_G1","Sex"))
dsD$DIF_G<-#abs
  (dsD$Intel_G1_S1-dsD$Intel_G1_S2)
dsD$DIF_A<-#abs
  (dsD$AFI_G1_S1-dsD$AFI_G1_S2)
max_val <- max(abs(dsD$DIF_G),na.rm=TRUE)
values <- seq(-max_val, max_val, length = length(shading))
coef<-coef(lm(AFI_G1 ~ Intel_G1, data = dsOutcomes))
max_val <- max(abs(dsD$DIF_G),na.rm=TRUE)

values <- seq(-max_val, max_val, length = length(shading))
g1<- ggplot(dsD, aes(y=AFI_G1_S1, x=Intel_G1_S1,colour=(DIF_G))) + 
  geom_point(size =.5,alpha = 1,na.rm=TRUE,position = position_jitter(w = 0.015, h = .75)) + geom_abline(intercept = coef[1], slope = coef[2])+
  scale_colour_gradientn(name="Sibling \nDifferences \nin Ability",colours=shading,values=values,rescaler = function(x, ...) x, oob = identity)+
  theme_min(size=8) +
     theme(plot.title=element_text(size=8),
   plot.margin=unit(c(0,0.0,0,0), "cm"),
   legend.position= "left",
   legend.key.width=unit(.5,"cm"))+ 
  ylab("Gen1 Age at First Intercourse") +
  xlab("Gen1 AFQT (Age 14-21)") + coord_cartesian(ylim=c(12,30))+
theme(text = element_text(size=8))#+
#ggtitle("Generation 1:\n Age at First Intercourse vs. Intelligence")

##### g12
dsLinking <- subset(Links79Pair, RelationshipPath=="Gen2Siblings")
dsOutcomes <- Gen2AFIConcor_3
